---
date: "2021-10-25T00:00:00Z"
external_link: ""
image:
#  caption: Photo by rawpixel on Unsplash
  focal_point: Smart
  
#slides: example
summary: How to write a user-defined function in R.
tags:
- Blogs
title: Write a user-defined function in R

categories: 
- R

output:
  blogdown::html_page:
    toc: true
  
---
## INTRODUCTION

In this tutorial, we will learn, how to make our own custom function in R.Though, R has thousands of functions under thousands of packages, but it is most important to know about how to make a customized function function.

## User Defined Functions

- Functions are created using the ***function()*** directive and are
stored as R objects just like anything else. In particular, they are R
objects of class function".

- The basic format of the code is 

**function_name = function(arguments)**   
**{**
  **main computation to be done**
**}**


```{r}
#---define a function
testfunction = function(x,y)
{
  x+y
}

#--- call the function with arguments 2,5
testfunction(2,5)
```

## Doing more than one computation

- When a function performs more than one task and gives multiple objects ***return()*** is used to get all the outputs in a form of a vector.


```{r}
testfunction = function(x,y)
{
  sum= x+y
  prod= x*y
  return(c(Sum=sum,Product=prod))
}
testfunction(2,5)
```

- Note that the two output can be accepted separatedly as 

```{r}
result = testfunction(2,5)
result[1]
```
```{r}
result[2]
```

- Alternatively multiple output can be extracted using ***list()***. This will enable us to extract by names (along with indices)

```{r}
testfunction = function(x,y)
{
  sum= x+y
  prod= x*y
  output=list(Sum=sum,Product=prod)
  return(output)
}
output= testfunction(2,5)
```


```{r}

output$Sum
```

```{r}
output$Product
```

## Default argument of a function

- R provides methods to define the default value of the arguments while defining the function.

- This default values will be used when the function is called unless this argument values are changed during calling.


```{r}
#--- initializing x=1 & y=1

testfunction = function(x=1,y=1)
{
  sum= x+y
  prod= x*y

  #--- Creates the output list
  
  output=list(Sum=sum,Product=prod)
  return(output)
}
testfunction() #-- calling function with no arguments
```

## Additional Arguments

- Provision for additional arguments (*probably optional arguments, which cannot be decided beforehand*) can be done using "**...**"

```{r}
testfunction = function(x=1,y=1,...)
{
  sum= x+y
  prod= x*y

  #--- Creates the output list
  
  output=list(Sum=sum,Product=prod)
  return(output)
}
testfunction(2,5,z=12) #-- z is an extra argument which has no use in this function
```

## Data types of arguments

- Since the types of arguments are not specified (*at the time of definition*), the arguments can be of any type of any data type provided the **internal code of the function is conformable with that data type**

```{r}
testfunction = function(x=1,y=1,...)
{
  sum= x+y
  prod= x*y

  #--- Creates the output list
  
  output=list(Sum=sum,Product=prod)
  return(output)
}
testfunction(2,5,z=12) #-- calling with vectors
```


```{r,eval=FALSE}
#-- calling with characters
testfunction("F","M")
```

**$\color{red}{\text{Error in x+y : non-numeric argument to binary operator}}$**


## Sanity checking argument

- So how can we stop a function if the user calls it with non-conformable arguments ?

- A good practice is to write functions in such that while calling, it checks whether the arguments supplied make sense before going to the main body of the function.


```{r,eval=FALSE}
testfunction = function(x=1,y=1,...)
{
  #-- check if the arguments are not characters
  stopifnot(typeof(x)!="character",typeof(y)!="character")
  sum= x+y
  prod= x*y

  #--- Creates the output list
  
  output=list(Sum=sum,Product=prod)
  return(output)
}
testfunction("F","M")
```

**$\color{red}{\text{Error in testfunction("F","M") : typeof(x) != "character" is not TRUE}}$**

- The **stopifnot** function halts the execution of the function (*with error message*) if all of its arguments do not evaluate to **TRUE**.

## Scope of variables

- When we define a **variable within a function**, it will be local and will not affect any **global variable** even if the name matches.

```{r,eval=FALSE}
f_outer=function()
{
  a=2
  f_inner=function()
  {
    b=5
  }
}
c=10
```

- Then variable **c** is global to both **f_outer** and **f_inner**. For **f_inner** variable **b** is local but **a** is global whereas for **f_outer**, both **a** and **b** are local.

## Recursive Function

- R supports recursive function, i.e., a function that calls itself recursively.

```{r}
#-- Creating a recursive function
fact= function(x)
{
  if(x==0)
  {
    return(1)
  }
  else
  {
    return(x+fact(x-1))
  }
}

fact(5) #-- calling the function with x=5
```

## Loops in R

- Loops helps to repeat a job. We first start with for loop.

- The syntax is
**for(variable in sequence)**
**{**
  **expression to be evaluated**
**}**

- Here sequence is an expression which evaluates to a vector(not necessarily in A.P.)

- For example all the following are valid
**for(i in 1:10)**
**for(i in c(2,3,7,9,13,17,19,23))**
**for(i in c("A","B","C"))**

- The no. of times the expression in loop is evaluated is the length of the sequence.

## While loop

- The syntax is 
**while(condition)**
**{**
  **expression to be evaluated**
**}**

- The loop repeats its action until the test condition is not satisfied.

- Unlike for loop we need not to know in advance how many times the loop will repeat.

## If & If-Else

- The syntax for if statement is
if(condition)
{
  expression
}

- For a binary situation we can use if-else
if(condition)
{
  expression 1
}
else
{
  expression 2
}

## If-Else function

- An alternative better way if-else statements is **ifelse()** function.

- The syntax is
**new variable= ifelse(Some Condition, Value of new variable if condition is true, value if condition is false)**

- e.g. 
**category= ifelse(marks>80, "Good","Fair")**
assigns value Good if marks is more than 80 and otherwise Fair.

- The additional advantage is in the condition this function can compare a vector with scalar (*interpreted as each element compared to the scalar*)

## Else if Ladder

- When we have more than two cases we can use else-if ladder

```{r,eval=FALSE}
f= function(x)
{
  if(x==1) print(a)
  else if(x==2) print(b)
  else print(c)
}
```

## Switch Statement

- An alternative and faster way is **switch()** statement.

- The basic syntax is **switch(statement,list)**

- Here **statement** is evaluated and based on this value, the corresponding item in the **list** is returned.

- e.g. **switch(2,"A","B","C")** gives the answer "B". It selects the item no. 2 from the list.

- **switch(4,"A","B","C")** gives NULL as there is no item with index 4 in the list.

- **switch("color","color"="red","shape"="round","length"=5)** gives answer red (*it matches the string*)


```{r}
stat= function(x,type)
{
  switch(type,"mean"=mean(x),
         "median"=median(x),
         "sd"=sd(x))
} #--- function ends here

stat(1:10,"mean") #-- call the function with mean
```

```{r}
stat(1:10,"median") #-- call the function with median
```

## Repeat Loop

- Basic syntax is 

repeat
{
  expression to be evaluated
}

- No default way of termination.

- We need to manually terminate the loop using **break** statement.

```{r}
x=1 #-- Take any value x as 1
repeat
{     #-- Loop begin here
  x=x+1
  if(x==6) break #-- manual instruction to exit loop
}   #-- Loop ends here

x #-- checking the value of x
```

## Plotting Functions

- Any function can be plotted using **curve()**

- The syntax is 
**curve(function,from,to,n,add=T/F,...)**
where from and to are range over which the function is plotted and **n**(*integer*) is the number of points at which we evaluate. **add=TRUE/FALSE** indicates whether to add this curve to a existing plot or not.

- To get more information about it's arguments type **??curve() **
  
```{r,eval=FALSE}
  myfun= function(x)
  {
    x*(1-x)
  }
  curve(myfun,from=0,to=1)
```
  

```{r,echo=FALSE}
  myfun= function(x)
  {
    x*(1-x)
  }
  curve(myfun,from=0,to=1)
```
  
## Plotting normal curve
    
  
```{r,eval=FALSE}
  #-- dnorm gives pdf of N(0,1)
  curve(dnorm,from = -4,to=4,n=500)
```
    
```{r,echo=FALSE}
  #-- dnorm gives pdf of N(0,1)
  curve(dnorm,from = -4,to=4,n=500)
```
  
## sin(1/x) plot
    
```{r}
  curve(sin(1/x),from = -2,to = 2)
```
  
## Zoom at the origin
  
```{r}
  curve(sin(1/x),from = -0.1,to = 0.1)
```
  
## Solving Equation
  
- Already we know if we have a system of equations we can use ***solve()***
    
- For equations involving one variable we can use ***uniroot()***
    
- The syntax is ***uniroot(function,interval,...)***
    
- For solve $$e^x=sin(x)$$ we write
  
```{r,eval=FALSE}
  uniroot(function(x) exp(x)-sin(x),c(-5,5))
```
  
## Solving Equation
  
```{r,echo=FALSE}
  uniroot(function(x) exp(x)-sin(x),c(-5,5))
```
  
## Solving Equation
  
- For finding real or complex roots of a ploynomial use ***polyroot()***
    
- For solving roots of $n$ non-linear equations we can use ***multiroot()*** from the ***rootSolve*** package.
  
## Some Calculus in R
  
- Define integral can be done using ***integrate()***
    
- e.g. $\int_0^1(x^2)dx$ can be done using
  
```{r}
  integrate(function(x) x^2,0,1)
```
  
- For derivatives, we use ***deriv()***
    
## Optimization
    
- Maximum or Minimum value of a function can be found using ***optimize()***
    
- ***optimize(function,interval,maximum=TRUE/FALSE)***
    
```{r}
  optimise(function(x) exp(-x),c(0,5))
```
  
- There are other functions for optimization like ***optim()***,***nlm()***,***constrOptim()***.
  

## Further reading

[*Garrett Grolemund*, **Hands-On Programming with R**, *O'REILLY*](https://rstudio-education.github.io/hopr/)