---
date: "2022-06-19T00:00:00Z"
external_link: ""
image:
#  caption: Photo by rawpixel on Unsplash
  focal_point: Smart
  
#slides: example
summary: The c3 package is a wrapper, or htmlwidget, for the C3 javascript charting library by Masayuki Tanaka. 

tags:
- Blogs
title: C3
categories:
- R
output:
  blogdown::html_page:
    toc: true
      
---
The `` c3 `` package is a wrapper, or htmlwidget, for the C3 javascript charting library by Masayuki Tanaka. You will find this package useful if you are wanting to create a chart using R and embedding it in a Rmarkdown document or Shiny App.

The ``C3`` library is very versatile and includes a lot of options. Currently this package wraps most of the ``C3`` options object. Even with this current limitation a wide range of options are available.


## Instalation

```{r,eval=FALSE}
install.packages("c3")

# or

devtools::install_github("mrjoh3/c3")

```

## Usage

The ``c3`` package is intended to be as simple and lightweight as possible. As a starting point the data input must be a ``data.frame`` or ``tibble`` with several options.

- If a ``data.frame`` without any options is passed all of the numeric columns will be plotted. This can be used in line and bar plots. Each column is a line or bar.

- For more complex plots only 3 columns are used, those defined as ``x``, ``y`` and ``group``. This requires a ``data.frame`` with a vertical structure.


### The Basics

Where no options are supplied a simple line plot is produced by default. Where no x-axis is defined the plots are sequential. ``Date`` x-axis can be parsed with not additional setting if in the format ``%Y-%m-%d`` (ie ‘2014-01-01’)

```{r}
library(c3)
data = data.frame(a = abs(rnorm(20) * 10),
                  b = abs(rnorm(20) * 10),
                  date = seq(as.Date("2011-01-01"), by = "month", length.out = 20))
c3(data)
```

### Piping

The package also imports the magrittr piping function ``(%>%)`` to simplify syntax.

```{r}
data%>%c3()
```

## Other Line Plots

There are 5 different line plots available:

- line
- spline
- step
- area
- area-step


### Spline

```{r}
data %>%
  c3() %>%
  c3_line('spline')
```


### Step

```{r}
data %>%
  c3(x = 'date') %>%
  c3_line('area-step')
```


## Bar Plots

```{r}
data[1:10, ] %>%
  c3() %>%
  c3_bar(stacked = TRUE, 
         rotate = TRUE)
```


## Mixed Geometry Plots

Mixed geometry currently only works with a horizontal ``data.frame`` where each numeric column is plotted.

```{r}
data$c <- abs(rnorm(20) *10)
data$d <- abs(rnorm(20) *10)
data %>%
  c3() %>%
  c3_mixedGeom(type = 'bar', 
               stacked = c('b','d'),
               types = list(a='area',
                            c='spline')
               )
```


## Secondary Y Axis

To use a secondary Y axis columns must first be matched to an axis and then the secondary axis made visible.

```{r}
data %>% 
  dplyr::select(date, a, b) %>%
  c3(x = 'date',
     axes = list(a = 'y',
                 b = 'y2')) %>% 
  c3_mixedGeom(types = list(a = 'line',
                            b = 'area')) %>% 
  y2Axis()
```

## Scatter Plot

```{r}
mtcars %>%
  c3(x = 'mpg', 
     y = 'wt', 
     group = 'cyl') %>% 
  c3_scatter()
```

## Pie Charts

```{r}
data.frame(India = 45,
           Bangladesh = 20,
           SriLanka = 10) %>% 
  c3() %>% 
  c3_pie()
```


## Donut Charts

```{r}
data.frame(red = 82, green = 33, blue = 93) %>% 
  c3(colors = list(red = 'red',
                   green = 'green',
                   blue = 'blue')) %>% 
  c3_donut(title = '#d053ee')
```

## Gauge Charts

```{r}
data.frame(data = 80) %>% 
  c3() %>% 
  c3_gauge()
```


## Grid Lines & Annotation

```{r}
data %>%
  c3() %>%
  grid('y') %>%
  grid('x', 
       show = F, 
       lines = data.frame(value = c(3, 10), 
                          text= c('Line 1','Line 2')))
```

## Sub-chart

```{r}
data %>%
  c3(x = 'date') %>%
  subchart()
```


## Color Palette

Plot color palettes can be changed to either ``RColorBrewer`` or ``viridis`` palettes using either ``RColorBrewer`` (S3 method) or ``c3_viridus``.

```{r}
data.frame(sugar = 20, 
           fat = 45, 
           salt = 10, 
           vegetables = 60) %>% 
  c3() %>% 
  c3_pie() %>%
  RColorBrewer()
```

```{r}
data.frame(sugar = 20, 
           fat = 45, 
           salt = 10, 
           vegetables = 60) %>% 
  c3() %>% 
  c3_pie() %>%
  c3_viridis()
```


## Point Size

```{r}
data %>%
  c3(x = 'date') %>%
  point_options(r = 6, 
                expand.r = 2)
```

## On Click

Onclick, onmouseover and onmouseout are all available via the ``c3`` function. To use wrap a js function as a character string to ``htmlwidgets::JS()``. Please see the ``C3.js`` documentation and examples. The example below should be enough to get you started.

```{r}
data %>% 
    c3(onclick = htmlwidgets::JS('function(d, element){console.log(d)}'))
```

## Tooltips

C3 tooltips are readily modified with the use of javascript functions. For further detail see the ``C3.js`` documentation. Or for more advanced usage see the ``C3.js`` examples page.

```{r}
library("htmlwidgets")
data %>%
  c3() %>%
  tooltip(format = list(title = JS("function (x) { return 'Data ' + x; }"),
                        name = JS('function (name, ratio, id, index) { return name; }'),
                        value = JS('function (value, ratio, id, index) { return ratio; }')))
```

